---
title: "Covid vaccine regression"
author: ""
date: "2022-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

## Regression problem

- We will run regression and other related models for Covid-19 vaccination data

## Libiraries

- We will use the following packages

```{r}
library(tidyverse)
library(caret)
library(glmnet)
```

## Load data

We will use the following data. It is a combined dataset from three data sources we have been using. The code for processing is available at `data_prep/data_preparation.R`.

```{r}
data_vac <- read_csv("data/vaccine-data.csv.gz") 
```


## Check data

Let's have a cursory look at the data, especially check the distribution of the output variable `Booster_Doses_18Plus_Vax_Pct` Do we need conversion?

### `head()`

```{r}
data_vac %>% head()
```

### Check the distribution of the output

```{r}
data_vac %>% ggplot(aes(x = Booster_Doses_18Plus_Vax_Pct)) + geom_density()
```


## Decide the variable to include as input

- There are 47 variables what are possible predictors? Especially:
  - pct_trump
  - demography: TotalPop, Men, Women, Hispanic, White, Black, Native, Asian, Pacific, VotingAgeCitizen, Income, IncomePerCap, Poverty, ChildPoverty, Professional, Service, Office, Construction, Production, Drive, Carpool, Transit, Walk, OtherTransp, WorkAtHome, MeanCommute, Employed, PrivateWork, PublicWork, SelfEmployed, FamilyWork, Unemployment
- What do you think should be included as the inputs?


```{r}
data_vac_use <- data_vac %>%
  select(Booster_Doses_18Plus_Vax_Pct, Black, Hispanic, IncomePerCap,
         Poverty, ChildPoverty, Unemployment, pct_trump) %>%
  drop_na()
```

## Data preparation

Here we need to prepare the data, in particular:

1. Train-test split
2. Data preprocessing

Using `caret` (or something else if you like), prepare two datasets of pre-processed train/test data.

## Train-test split

```{r}
set.seed(20220917)
train_id <- createDataPartition(data_vac_use$Booster_Doses_18Plus_Vax_Pct, times = 1, 
                    p = .7, list = F)
df_train <- data_vac_use %>% slice(train_id) 
df_test <- data_vac_use %>% slice(-train_id) 
#df_train <- data_vac_use[train_id, ]

# for (i in 1:10){
#   set.seed(2022)
#   set.seed(NULL)
#   sample(10,2) %>% print()
# }

```

## Preprocess

```{r}
prep <- df_train %>% 
  select(-Booster_Doses_18Plus_Vax_Pct) %>% 
  #select(-dummy1:dummy10)
  preProcess(method = c('center', 'scale'))
prep
df_train_prepped <- df_train %>% predict(prep, .)
df_test_prepped <- df_test %>% predict(prep, .)


```


## Analysis

### Linear regression

- Run linear regression 
- Evaluate the model

```{r}
mod_lm <- lm(Booster_Doses_18Plus_Vax_Pct ~ ., data = df_train_prepped)
summary(mod_lm)

```
```{r}
rmse <- function(errors){
  return((errors^2) %>% mean() %>% sqrt())
}

pred_train <- predict(mod_lm)
error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_train)

pred_test <- predict(mod_lm, newdata = df_test_prepped)
error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_test)


```

### Additional movel evaluations

Using the linear regression model as the baseline we attempt two things:

1. Is it possible to improve the prediction using more flexible models?
  - KNN-regression
  - Or regression model variant of models covered in classificaiton section. 
    - For example:
      - svm: svmPoly, svmRadial works both regression and classification (svmPoly may take quite long time as the number of tuning paramters are many.)
      - trees: rf
      


```{r}
ctrl <- trainControl(method = 'repeatedcv', repeats = 3, number = 5)
mod_knn <- train(Booster_Doses_18Plus_Vax_Pct ~ ., 
                 method = 'knn',
                 data = df_train_prepped, 
                 trControl = ctrl, 
                 tuneGrid = expand.grid(k = c(1:10, 20, 30)))
mod_knn

pred_train <- predict(mod_knn)
error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_train)

pred_test <- predict(mod_knn, newdata = df_test_prepped)
error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_test)



```

### SVM with Radial Kernel

```{r}
ctrl <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 1)
mod_svmradial <- train(Booster_Doses_18Plus_Vax_Pct ~ ., data = df_train_prepped, 
   method = "svmRadial", trControl = ctrl) 

mod_svmradial


pred_train <- predict(mod_svmradial)
error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_train)

pred_test <- predict(mod_svmradial, newdata = df_test_prepped)
error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_test)

```
```{r}
mod_rf <- train(Booster_Doses_18Plus_Vax_Pct ~ ., data = df_train_prepped, 
   method = "rf", trControl = ctrl) 

mod_rf


# pred_train <- predict(mod_svmradial)
# error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
# rmse(error_train)
# 
# pred_test <- predict(mod_svmradial, newdata = df_test_prepped)
# error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
# rmse(error_test)


```


## LASSO and ridge regression

- Now, let's run LASSO and/or Ridge regression. 
- What do you find? 
  - Shrinkage of the coefficients

### LASSO Outcome

```{r}
mat_train_x <- df_train_prepped %>% select(-Booster_Doses_18Plus_Vax_Pct) %>% as.matrix()
mat_test_x <- df_test_prepped %>% select(-Booster_Doses_18Plus_Vax_Pct) %>% as.matrix()

```


```{r}
mod_lasso <- cv.glmnet(mat_train_x, df_train_prepped$Booster_Doses_18Plus_Vax_Pct, 
          alpha = 1, 
          type.measure = 'mse',
          family = "gaussian")

coef(mod_lasso)
plot(mod_lasso)
plot(mod_lasso$glmnet.fit, xvar = "lambda")
#summary(mod_lasso$glmnet.fit)

pred_train <- predict(mod_lasso, newx = mat_train_x)
error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_train)

pred_test <- predict(mod_lasso, newx = mat_test_x)
error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_test)

```


```{r}
mod_ridge <- cv.glmnet(mat_train_x, df_train_prepped$Booster_Doses_18Plus_Vax_Pct, 
          alpha = 0, 
          type.measure = 'mse',
          family = "gaussian")

coef(mod_ridge)
plot(mod_ridge)
plot(mod_ridge$glmnet.fit, xvar = "lambda")
#summary(mod_ridge$glmnet.fit)

pred_train <- predict(mod_ridge, newx = mat_train_x)
error_train <- pred_train - df_train_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_train)

pred_test <- predict(mod_ridge, newx = mat_test_x)
error_test <- pred_test - df_test_prepped$Booster_Doses_18Plus_Vax_Pct
rmse(error_test)

```

#### Plot with `plot_glmnet`

Shrinkage plot of `glmnet` is not informative as it won't show the variable name. Instead you can use `plot_glmnet` in `plotmo` package.

```{r}
plotmo::plot_glmnet(mod_lasso$glmnet.fit, xvar = "lambda")
plotmo::plot_glmnet(mod_ridge$glmnet.fit, xvar = "lambda")
```




### Compare coefs: lm, lasso/ridge

Compare the cefficients across the models. What do you find?

```{r}
list(mod_lm, mod_lasso, mod_ridge) %>%
  map(~coef(.) %>% as.matrix %>% as.data.frame) %>% bind_cols() %>%
  rename(lm = 1, lasso = 2, ridge = 3)

list(mod_lm, mod_lasso, mod_ridge) %>%
  map(~coef(.) %>% as.matrix %>% as.data.frame) %>% bind_cols() %>%
  rename(lm = 1, lasso = 2, ridge = 3) %>%
  rownames_to_column(var = "variable") %>% pivot_longer(2:4) %>%
  filter(variable != "(Intercept)") %>%
  ggplot(aes(x = value, y = variable, color = name)) + geom_point() +
  theme_minimal()

```